---
title: NBFIs and the Stress in Money Markets during the Covid-crisis
subtitle:  Topic 8
author: Finn-Ole Höner
institute: Seminar on Non-Bank Financial Intermediaries (Dr. Andreas Schrimpf)
date: "22-01-2022"
fontsize: 12pt
output:
  beamer_presentation:
#    keep_tex: true
    toc: false
    slide_level: 2
    includes:
      in_header: tex/beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: citations.bib 

---
```{r, echo=FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(kableExtra)
library(vistime)
library(grid)
library(gridExtra)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, out.width = "85%", out.height = "100%", dpi=300)
options(knitr.table.format = "latex")
fs <- theme(text = element_text(size = 18), axis.text = element_text(size = 14))
```

# The March Market Turmoil

## The Covid-19 Pandemic

```{r fig.align='center'}
plot_covid + labs(caption = "", x="")  + fs
```

## Disturbances in Financial Markets

```{r covid_fm, fig.align='center'}
gridExtra::grid.arrange(plot_msci + labs(title = "", caption = "", x="") + fs, plot_yield_spread + labs(caption = "", x="") + fs, nrow = 2, top = textGrob("MSCI World and 10y Government Bonds",gp=gpar(fontsize=20,font=3)))
```

## The Money-Market and its NBFIs

Short term funding market, which provides funding to banks, corporations and governments. Borrowers look for financing of operations, investors look to invest their money at a market rate of return.

- Money Market Funds
- Broker Dealers
- Hedge Funds
- Principal Trading Firms

\bigskip

### Focus on three Key Players in the US market

- Government and Prime MMFs
- Broker Dealers
- *Central banks (the Fed)*

## Differences between Government and Prime MMFs
::: rows

:::: row

\centering
```{r, fig.align='center'}
df_table %>% kable(align = "c")
```

::::
\vfill
:::: row

```{r, fig.align='center', out.height="60%", out.width="50%"}
plot_orf_invest_type + labs(caption = "", x = "") + fs
```

::::

:::


# Analysis of these actors in the March Market Turmoil

## Research Questions

1. *What was the role of NBFIs in the market stress of March 2020? Which factors related to the business models of NBFIs played a destabilizing role?*

2. *What role did broker-dealers play during the stress episode?*

3. *What was the role of central banks and through which tools did they intervene to restore market functioning? What tools were effective and why?*

## RQ1 - Role of the NBFIs

- Flow from prime to government MMFs [see @li2021; @eren2020a; @fsb2020a]
- Broker dealers tried to intermediate fire-sales [see @fsb2020a]
- Prime MMFs liquidity supply dried up, increasing funding costs for others

## RQ1 - Role of the NBFIs

```{r libor, fig.align='center'}

grid.arrange(plot_libor_tenor + labs(x = "", caption = "") + fs + theme(legend.position = "bottom") + guides(colour = guide_legend(nrow = 1)), plot_libor_mat + labs(x = "", caption = "") + fs + theme(legend.position = "none"), nrow = 2, top = textGrob("LIBOR",gp=gpar(fontsize=20,font=3))) 

```


## RQ1 - NBFIs' Destabilizing Factors

@bouveret2021 collect a couple of destabilizing factors:

- MMF's large footprint in e.g. the repo market lead to disruptions [also @paddrik2021]
- Portfolio overlap 
- Low liquidity
- Floating NAV
- Run risk (redemption gates and fees, credit risk)


## RQ2 - Role of Broker Dealers
- Difficulties to take on new liabilities (deleveraging) and to sell off acquired assets (lack of cash) [compare @aramonte2021]
- Balance sheet constraints
- Increased funding costs (repos became more expensive)
- Operational struggles (move to home-office) [@fsb2021]
- Insertion point for central bank interventions [@boyarchenko2021]

## RQ3 - Central Bank Interventions
::: rows

:::: row

The Fed intervened with a multitude of measures targeted at different areas of the financial markets.

\vfill

::::

:::: row
\centering
```{r, fig.align="center"}
cb_table %>% knitr::kable(align = "c")
```

::::
:::

## RQ3 - Tools' Effectiveness
- Existing SEC regulations (gates and fees) were detrimental [@li2021]
- Rebound flows to prime funds, due to loans for broker dealers (MMLF)
- Market improvements for CP following announcement of the CPFF
- Foreign exchange swap lines helped to divert some stress away from US MMFs and to transmit the Fed's interventions


# Summary - RQs

1) The fragility of prime MMFs led to their liquidity provision coming to a halt, causing problems for other market participants
2) Broker dealers were not able to intermediate the large, one-directional, flow of assets
3) Central bank interventions were key to remove liquidity frictions, especially the MMLF

## References {.allowframebreaks}

